---
title: "project-winter-update"
author: "H. Hunter"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:  
  bookdown::html_document2:
    collapsed: no
    fig_caption: yes
    highlight: espresso
    number_sections: yes
    smooth_scroll: yes
    theme: sandstone
    toc: yes
    toc_float: yes
---
```{=html}
<style type="text/css">

body{ /* Normal  */
      font-size: 13px;
  }
td {  /* Table  */
  font-size: 13px;
}
h1.title {
  font-size: 24px;
  color: Black;
}
h1 { /* Header 1 */
  font-size: 19px;
  color: Black;
}
h2 { /* Header 2 */
  font-size: 15px;
  color: Black;
}
h3 { /* Header 3 */
  font-size: 15px;
  color: Black;
}
</style>
```

```{r setup, include=FALSE}

# global settings
knitr::opts_chunk$set(echo = FALSE, message=FALSE, cache=FALSE, include=TRUE, dpi=300)
options(scipen=1, digits=4)

# functions
source("../functions.R")

# libraries
source("../load.R")
library(plotly)

# read brood table data
sock <- read.csv("../data/sockeye/master_sockeye_brood_table_covar.csv", stringsAsFactors = F)
pink <- read.csv("../data/pink/master_pink_brood_table_covar_0124.csv", stringsAsFactors = F)
chum <- read.csv("../data/chum/master_chum_brood_table_covar.csv", stringsAsFactors = F)
# read info tables
sock.info <- read.csv("../data/sockeye/master_sockeye_stock_info.csv", stringsAsFactors = F)
pink.info <- read.csv("../data/pink/master_pink_stock_info_0124.csv", stringsAsFactors = F)
chum.info <- read.csv("../data/chum/master_chum_stock_info.csv", stringsAsFactors = F)

# subset sockeye stocks
stk.sub <- sock.info$Stock.ID[sock.info$yr_start < 1985 & sock.info$yr_end >= 2014]
sock <- sock %>% filter(Stock.ID %in% stk.sub)
sock.info <- sock.info %>% filter(Stock.ID %in% stk.sub)

# everything crumbles if they aren't ordered factors
sock <- geographic.order(sock)
sock.info <- geographic.order(sock.info)
pink <- geographic.order(pink)
pink.info <- geographic.order(pink.info)
chum <- geographic.order(chum)
chum.info <- geographic.order(chum.info)

# remove the data_master and data_info that functions rely on
rm("data_master")
rm("info_master")


# figure specifications
bayesplot::bayesplot_theme_set(new = theme_sleek())
col.region <- rev(chroma::qpal(7, luminance = 40)[c(1, 3, 5, 7)]) # region colours
col.dk <- rev(chroma::qpal(7, luminance = 20)[c(1, 3, 5, 7)]) # darker region colours
names(col.region) <- names(col.dk) <- unique(sock.info$ocean_region_lab) #name
shp.reg <- c(18, 16, 17, 15) # region shapes
names(shp.reg) <- unique(sock.info$ocean_region_lab) # name
sp.col <- c("seagreen4", "palevioletred3", "orangered4") # species colours 
names(sp.col) <- c("Chum", "Pink", "Sockeye") # name

```

# Purpose

This document summarizes updated results from the analyses presented [here](https://pacific-salmon-assess.github.io/Salmon-warming-crowded-oceans/Rmd/project-fall-update.html). Notably, the analysis now includes Pink and Chum salmon stocks (N = 67, N = 48) from across the eastern North Pacific, as well as newly added Sockeye stocks (19 new stocks for a total of N = 55). 


# Methods

The overarching goal of our analyses is to characterize relationships between ocean conditions (SST and abundance of potential competitors) and salmon productivity (R/S) over space and time. Since our [last update](https://pacific-salmon-assess.github.io/Salmon-warming-crowded-oceans/Rmd/project-fall-update.html#2_Methods), we have made minor changes to our approach, as described in the sections below.

## Data

We have compiled spawner-recruitment time series for all five species of Pacific salmon - [available here](https://github.com/Pacific-salmon-assess/dfo_salmon_compilation), although only Sockeye, Pink, and Chum populations are analyzed here. We analyze a total of 170 populations originating from Washington to the Bering Sea:

* From our original set of 58 Sockeye time series, we obtained more data from the SEAK and WC regions, but subsequently chose to remove time series that did not span at least the brood years of 1984-2015 without significant gaps, ensuring sufficient overlap with each time period in our era models and a suitable length for random walk models. There are now 55 Sockeye stocks in the analysis - many of the original stocks, and some new ones.

* We obtained spawner-recruitment time series for Pink and Chum populations, and currently use all time series with good data quality, regardless of whether they meet the criteria applied to Sockeye data. These data do not represent all regions and time periods well, and notably, we have no data after 2009 for all Alaskan stocks. We are working to update these data through to the present.

* Compilation of coho and Chinook time series is ongoing.


### Map of ocean entry points

The map below shows the ocean entry locations of all stocks in the analysis presented. The legend to the right of the plot allows you to toggle between species (double-click a legend item to isolate that species). Hovering over a point will show the stock name and ocean region.

```{r Ocean entry map, out.width="100%", out.height="50%", warning=FALSE}
# specify output size or plotly makes it huge

# Downlad map and convert to sp
na_map <- rnaturalearth::ne_states(country = c("United States of America", "Canada"), returnclass='sf')

axes <- list( xlims=c(-171, -121), 
              ylims=c(46, 67),
              xbreaks=seq(-170,-120,10), 
              xlabels=as.character(seq(-170,-120,10)),
              seq(45, 65, 5), 
              ybreaks=seq(45, 65, 5),
              ylabels=as.character(seq(45,65,5)))

#make map data
sock.info$Species <- "Sockeye"
chum.info$Species <- "Chum"
pink.info$Species <- "Pink"
map.info <- rbind(sock.info, chum.info, pink.info)


map <- ggplot(data = map.info) + 
  geom_sf(data=na_map, color="grey40", fill="white", linewidth=0.1) + 
  ggspatial::geom_spatial_point(aes(x=lon, y=lat, col=Species, shape=Species, group=Stock, text=paste("Region:", ocean_region_lab)), 
                                crs=4326, size=2.5,
                                position=position_jitter(w=0.2, h=0.2)) +
  coord_sf(xlim=axes$xlims, ylim=axes$ylims) +
  scale_x_continuous(breaks=axes$xbreaks, labels=axes$xlabels) +
  scale_y_continuous(breaks=axes$ybreaks, labels=axes$ylabels) +
  scale_colour_manual(values=sp.col) +
  scale_shape_manual(values=c(15,16,17)) +
  labs(x="Longitude (°E)", y="Latitude (°N)") +
  theme_sleek() + 
  theme(panel.grid = element_blank(),
        plot.title = element_text(hjust=0.5),
        legend.position = c(0.4,0.25),
        legend.background = element_rect(colour="grey75")
  )

ggplotly(map, tooltip=c("group", "text")) %>% config(modeBarButtonsToRemove = 
                                                       c("zoom", "lasso2d", "select2d", 
                                                         "autoscale2d", "hoverClosestCartesian", "hoverCompareCartesian")
                                                     #,scrollZoom = TRUE
                                                     )
#%>% layout(legend=list(x=0.8,y=0.9))

```
**Figure 1**. Ocean entry locations of Sockeye (n=55), Pink (n=67), and Chum (n=48) salmon stocks included in analyses (total N=170).


### Time Series Length {.tabset}

Productivity (log[R/S]) time series of all stocks are shown below to illustrate the length and relative number of time series among regions and species. Vertical dashed lines represent proposed regime shifts in the 'era' model (described in sec. 2.2) to illustrate the coverage of data over periods of interest.

#### Sockeye

```{r Sockeye time series, fig.height=6.5}
prod_dat <- ocean_region_lab(sock)
prod_dat <- stock.plot.lab(prod_dat, numbered = F)

ggplot(prod_dat) + 
  geom_vline(xintercept=c(1989,2011), color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_line(aes(x=BY, y=lnRS, col=ocean_region_lab)) + 
  facet_grid(rows=vars(stock_lab), switch ="y", scales="free_y", as.table=F) + 
  scale_colour_manual(values=col.region) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y.left = element_text(angle=0, hjust=0, margin=margin(l=0, r=0)),
        strip.background = element_rect(fill="transparent", colour="transparent"),
        strip.text = element_text(size=7, ),
        panel.spacing.y = unit(0, unit="cm"),
        panel.background = element_rect(fill="white"),
        legend.position = "none") +
  labs(y="", x="Brood Year")


```
**Figure 2**. Productivity (log[R/S]) time series of Sockeye stocks (n=55) with vertical dashed lines indicating proposed ocean regime shifts.

#### Pink

```{r Pink time series}
prod_dat <- ocean_region_lab(pink)
prod_dat <- stock.plot.lab(prod_dat, numbered = F)

# should also show gaps!
# is there a way to show which series are included in each model type?
# figure out whether trimmed era model should be used...
# if so show the trim

ggplot(prod_dat) + 
  geom_vline(xintercept=c(1989,2011), color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_line(aes(x=BY, y=lnRS, col=ocean_region_lab)) + 
  facet_grid(rows=vars(stock_lab), switch ="y", scales="free_y", as.table=F) + 
  scale_colour_manual(values=col.region) +
  scale_x_continuous(breaks=seq(1950, 2020, 10)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y.left = element_text(angle=0, hjust=0, margin=margin(l=0, r=0)),
        strip.background = element_rect(fill="transparent", colour="transparent"),
        strip.text = element_text(size=7, ),
        panel.spacing.y = unit(0, unit="cm"),
        panel.background = element_rect(fill="white"),
        legend.position = "none") +
  labs(y="", x="Brood Year")

```
**Figure 3**. Productivity (log[R/S]) time series of Pink stocks (n=67) with vertical dashed lines indicating proposed ocean regime shifts. Note that Even- and Odd-run pinks are considered separately due to their independent and at times vastly different productivity regimes.

#### Chum 


```{r Chum time series}
prod_dat <- ocean_region_lab(chum)
prod_dat <- stock.plot.lab(prod_dat, numbered = F)

ggplot(prod_dat) + 
  geom_vline(xintercept=c(1989,2011), color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_line(aes(x=BY, y=lnRS, col=ocean_region_lab)) + 
  facet_grid(rows=vars(stock_lab), switch ="y", scales="free_y", as.table=F) + 
  scale_colour_manual(values=col.region) +
  scale_x_continuous(breaks=seq(1950, 2020, 10)) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.y.left = element_text(angle=0, hjust=0, margin=margin(l=0, r=0)),
        strip.background = element_rect(fill="transparent", colour="transparent"),
        strip.text = element_text(size=7, ),
        panel.spacing.y = unit(0, unit="cm"),
        panel.background = element_rect(fill="white"),
        legend.position = "none") +
  labs(y="", x="Brood Year")

```
**Figure 4**. Productivity (log[R/S]) time series of Chum stocks (n=48) with vertical dashed lines indicating proposed ocean regime shifts.

## Models

Our updated analyses focus on three classes of generalized spawner-recruitment models:

1. Stationary models (e.g. [Connors et al. 2020](https://cdnsciencepub.com/doi/full/10.1139/cjfas-2019-0422)), which estimate time-invariant relationships. These have not changed since last update other than the addition of data for the two new species.

2. 'Era' models (e.g., [Malick 2020](https://onlinelibrary.wiley.com/doi/10.1111/fog.12469)), which allow relationships to vary among pre-defined periods that represent hypothesized shifts in North Pacific Ocean processes and relationships. 

* We no longer model 1976/77 as a regime shift as per feedback from our last update, which highlighted that this shift from a cold to warm PDO phase simply reflected a temperature change (which is already accounted for via our direct consideration of ocean temperature) and not a fundamental change in oceanographic processes like the 1988/89 regime shift.

* Our updated analyses consider a potential regime shift at the onset of the ~2013-2015 marine heatwave ('the blob') [sources?]. Currently, we consider brood years >= 2011 to have interacted with the marine heatwave regardless of the species and stock. In future analyses, the timing can be altered to better reflect the diverse life histories of populations in the analysis.

3. Random walk models (e.g., [Malick 2020](https://onlinelibrary.wiley.com/doi/10.1111/fog.12469)), which allow relationships to evolve gradually through time. 

* These have not changed since last update other than the addition of new data. However, a version that estimates ocean basin-scale trends in relationships (e.g., as shared trends within ocean region that are modeled hierarchically) is currently in initial stages of application.

(4) At this time we are not pursuing Hidden Markov models, which model the system as oscillating between two or more latent states, as they did not capture trends shown in both other models. They may be revisited in the future.


Details on each model class and our Bayesian model fitting procedure can be found [here](https://pacific-salmon-assess.github.io/Salmon-warming-crowded-oceans/Rmd/project-fall-update.html#23_Era_models).


# Results

## Overall insights

To be discussed Feb 27 ***** 

Inference
- We see evidence of nonstationarity across all 3 species
- The degree of change varies among regions and species
- There is some evidence that the 'marine heatwave' period changed ocean condition-productivity relationships


## Stationary models {.tabset}


### Sockeye

```{r Sockeye stationary}

# Load model fit
load("../output/models/stat/sockeye/stat_sub.Rdata")

# Get data
lst <- hb05_density_df(stat_sub, ocean.regions = 4, info_master=sock.info, data_master=sock)
s.df <- lst$stock
m.df <- lst$region
m.df$region <- factor(m.df$region, levels = c("West Coast", "Gulf of Alaska", "Southeast Alaska", "Bering Sea"))

## Covariate labels
vars <- data.frame(var = levels(m.df$var))
vars$lab <- gsub("Comp", "Competitors", vars$var)
vars$var <- factor(vars$var, levels = c("SST", "Comp", "SST + Comp"))

# I think I could shorten this 
ggplot(m.df) +
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(data = s.df[s.df$region == "West Coast", ],
            aes(x = x, y = y, group = stock), color = col.region[["West Coast"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Southeast Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Southeast Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Gulf of Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Gulf of Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Bering Sea", ],
            aes(x = x, y = y, group = stock), color = col.region[["Bering Sea"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(aes(x = x, y = y, color = region), linewidth = 1, alpha=1, 
            na.rm = TRUE) +
  scale_colour_manual(name = "Ocean Region", values=col.region, guide="legend") +
  labs(x = "Percent change in R/S",
       y = "Posterior density",
       color = "") +
  scale_x_continuous(limits = c(-50, 50), expand = c(0, 0)) +
  scale_y_continuous(breaks=NULL) +
  geom_text(data = vars,
            aes(x = -48.1,
                y = max(m.df$y) - 0.008,
                label = lab),
            hjust = 0,
            size = 2.7,
            color = "grey30") +
  facet_wrap( ~ var, ncol = 1) +
  theme_sleek(base_size = 9) +
  theme(legend.justification = c(0, 0),
        legend.position = c(0.83, 0.4),
        legend.key.size = unit(10, "pt"),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        panel.spacing.y = unit(-0.5, "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())


```
**Figure 5**. Posterior probability distributions of the predicted effect of SST (top), competitor abundance (middle), and the combined effect (bottom) on Sockeye productivity (R/S). Faint lines show stock-specific effects while bold lines show regional effects from hierarchical model. X-axis values represent the percent change in productivity per standard deviation unit increase in the covariate.

### Pink
```{r Pink stationary}

# Load model fit
load("../output/models/stat/pink/stat_a.Rdata")

# Get data
lst <- hb05_density_df(stat_a, ocean.regions = 4, info_master=pink.info, data_master=pink)
s.df <- lst$stock
m.df <- lst$region
m.df$region <- factor(m.df$region, levels = c("West Coast", "Gulf of Alaska", "Southeast Alaska", "Bering Sea"))

## Covariate labels
vars <- data.frame(var = levels(m.df$var))
vars$lab <- gsub("Comp", "Competitors", vars$var)
vars$var <- factor(vars$var, levels = c("SST", "Comp", "SST + Comp"))

# I think I could shorten this sometime
ggplot(m.df) +
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(data = s.df[s.df$region == "West Coast", ],
            aes(x = x, y = y, group = stock), color = col.region[["West Coast"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Southeast Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Southeast Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Gulf of Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Gulf of Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Bering Sea", ],
            aes(x = x, y = y, group = stock), color = col.region[["Bering Sea"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(aes(x = x, y = y, color = region), linewidth = 1, alpha=1, 
            na.rm = TRUE) +
  scale_colour_manual(name = "Ocean Region", values=col.region, guide="legend") +
  labs(x = "Percent change in R/S",
       y = "Posterior density",
       color = "") +
  scale_x_continuous(limits = c(-50, 100), expand = c(0, 0)) +
  scale_y_continuous(breaks=NULL) +
  geom_text(data = vars,
            aes(x = -48.1,
                y = max(m.df$y) - 0.008,
                label = lab),
            hjust = 0,
            size = 2.7,
            color = "grey30") +
  facet_wrap( ~ var, ncol = 1) +
  theme_sleek(base_size = 9) +
  theme(legend.justification = c(0, 0),
        legend.position = c(0.83, 0.4),
        legend.key.size = unit(10, "pt"),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        panel.spacing.y = unit(-0.5, "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())
        
```
**Figure 6**. Posterior probability distributions of the predicted effect of SST (top), competitor abundance (middle), and the combined effect (bottom) on Pink salmon productivity (R/S). Faint lines show stock-specific effects while bold lines show regional effects from hierarchical model. X-axis values represent the percent change in productivity per standard deviation unit increase in the covariate.

### Chum

```{r Chum stationary}

# Load model fit
load("../output/models/stat/chum/stat_a.Rdata")

# Get data
lst <- hb05_density_df(stat_a, ocean.regions = 4, info_master=chum.info, data_master=chum)
s.df <- lst$stock
m.df <- lst$region
m.df$region <- factor(m.df$region, levels = c("West Coast", "Gulf of Alaska", "Southeast Alaska", "Bering Sea"))

## Covariate labels
vars <- data.frame(var = levels(m.df$var))
vars$lab <- gsub("Comp", "Competitors", vars$var)
vars$var <- factor(vars$var, levels = c("SST", "Comp", "SST + Comp"))

# I think I could shorten this sometime
ggplot(m.df) +
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(data = s.df[s.df$region == "West Coast", ],
            aes(x = x, y = y, group = stock), color = col.region[["West Coast"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Southeast Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Southeast Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Gulf of Alaska", ],
            aes(x = x, y = y, group = stock), color = col.region[["Gulf of Alaska"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(data = s.df[s.df$region == "Bering Sea", ],
            aes(x = x, y = y, group = stock), color = col.region[["Bering Sea"]], alpha=0.3,
            na.rm = TRUE) +
  geom_path(aes(x = x, y = y, color = region), linewidth = 1, alpha=1, 
            na.rm = TRUE) +
  scale_colour_manual(name = "Ocean Region", values=col.region, guide="legend") +
  labs(x = "Percent change in R/S",
       y = "Posterior density",
       color = "") +
  scale_x_continuous(limits = c(-50, 50), expand = c(0, 0)) +
  scale_y_continuous(breaks=NULL) +
  geom_text(data = vars,
            aes(x = -48.1,
                y = max(m.df$y) - 0.008,
                label = lab),
            hjust = 0,
            size = 2.7,
            color = "grey30") +
  facet_wrap( ~ var, ncol = 1) +
  theme_sleek(base_size = 9) +
  theme(legend.justification = c(0, 0),
        legend.position = c(0.83, 0.4),
        legend.key.size = unit(10, "pt"),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        panel.spacing.y = unit(-0.5, "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())

```
**Figure 7**. Posterior probability distributions of the predicted effect of SST (top), competitor abundance (middle), and the combined effect (bottom) on Chum salmon productivity (R/S). Faint lines show stock-specific effects while bold lines show regional effects from hierarchical model. X-axis values represent the percent change in productivity per standard deviation unit increase in the covariate.


## Era models {.tabset}

### Sockeye
```{r Sockeye eras}
load("../output/models/dyn/sockeye/hbm_era_2c_3sub.RData")

## Data
# Density dataframe - by stock
dens.df.st.2c <- era_density_df(era.2c.3sub, par=c("gamma", "kappa"), percent.change = T, neras=3, info=sock.info)
dens.df.st.2c <- ocean_region_lab(dens.df.st.2c)
dens.df.st.2c <- era_lab(dens.df.st.2c, data_master=sock, breakpoint2=2011)

# Density dataframe (regional lvl)
dens.df.reg.2c <- era_density_df(era.2c.3sub, par=c("gamma", "kappa"), mu=T, percent.change=T, neras=3)
dens.df.reg.2c <- ocean_region_lab(dens.df.reg.2c)
dens.df.reg.2c <- era_lab(dens.df.reg.2c, data_master=sock, breakpoint2=2011)



## Figures
ggplot(dens.df.st.2c) + 
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(aes(x=x, y=dens, group=stock, col=ocean_region_lab), alpha=0.2) + 
  geom_path(data=dens.df.reg.2c, aes(x=x, y=dens, col=ocean_region_lab), alpha=0.85, linewidth=1) +
  scale_colour_manual(values=col.region) +
  facet_grid(rows=vars(era_lab), cols=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  coord_cartesian(xlim=c(-50,50)) +
  theme_sleek() + labs(x="Percent change in R/S", y="", col="Ocean Region") +
  theme(axis.text.y=element_blank(),
        legend.key.size = unit(10, "pt"),
        legend.position=c(0.9, 0.2),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size=9),
        panel.spacing.y = unit(-0.5, "pt"),
        axis.ticks.y=element_blank()) 


```
**Figure 8**. Posterior probability distributions of the predicted effect of SST and competitors on Sockeye productivity over three pre-defined time periods/eras (earliest in top panel). Regional mean effects are shown by bold lines and individual stocks’ distributions by light lines.


### Pink
```{r Pink eras}

load("../output/models/dyn/pink/hbm_era_2c_3.RData")

## Data
# Density dataframe - by stock
dens.df.st.2c <- era_density_df(era.2c.3, par=c("gamma", "kappa"), percent.change = T, neras=3, info=pink.info)
dens.df.st.2c <- ocean_region_lab(dens.df.st.2c)
dens.df.st.2c <- era_lab(dens.df.st.2c, data_master=pink, breakpoint2=2011)
dens.df.st.2c <- dplyr::filter(dens.df.st.2c, !(Ocean.Region2 %in% c("BS", "GOA") & era=="Late"))

# Density dataframe (regional lvl)
dens.df.reg.2c <- era_density_df(era.2c.3, par=c("gamma", "kappa"), mu=T, percent.change=T, neras=3, info=pink.info)
dens.df.reg.2c <- ocean_region_lab(dens.df.reg.2c)
dens.df.reg.2c <- era_lab(dens.df.reg.2c, data_master=pink, breakpoint2=2011)
dens.df.reg.2c <- dplyr::filter(dens.df.reg.2c, !(Ocean.Region2 %in% c("BS", "GOA") & era=="Late"))



## Figures
ggplot(dens.df.st.2c) + 
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(aes(x=x, y=dens, group=stock, col=ocean_region_lab), alpha=0.2) + 
  geom_path(data=dens.df.reg.2c, aes(x=x, y=dens, col=ocean_region_lab), alpha=0.85, linewidth=1) +
  scale_colour_manual(values=col.region) +
  facet_grid(rows=vars(era_lab), cols=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  coord_cartesian(xlim=c(-50,100)) +
  theme_sleek() + labs(x="Percent change in R/S", y="", col="Ocean Region") +
  theme(axis.text.y=element_blank(),
        legend.key.size = unit(10, "pt"),
        legend.position=c(0.9, 0.2),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size=9),
        panel.spacing.y = unit(-0.5, "pt"),
        axis.ticks.y=element_blank()) 

```
**Figure 9**. Posterior probability distributions of the predicted effect of SST and competitors on Pink salmon productivity over three pre-defined time periods/eras (earliest in top panel). Regional mean effects are shown by bold lines and individual stocks’ distributions by light lines. Note that data from Alaska for the most recent time period are needed (Figure 3).

### Chum
```{r Chum eras}
load("../output/models/dyn/chum/hbm_era_2c_3.RData")


## Data
# Density dataframe - by stock
dens.df.st.2c <- era_density_df(era.2c.3, par=c("gamma", "kappa"), percent.change = T, neras=3, info=chum.info)
dens.df.st.2c <- ocean_region_lab(dens.df.st.2c)
dens.df.st.2c <- era_lab(dens.df.st.2c, data_master=chum, breakpoint2=2011)
dens.df.st.2c <- dplyr::filter(dens.df.st.2c, !(Ocean.Region2 %in% c("BS", "GOA", "SEAK") & era=="Late"))

# Density dataframe (regional lvl)
dens.df.reg.2c <- era_density_df(era.2c.3, par=c("gamma", "kappa"), mu=T, percent.change=T, neras=3, info=chum.info)
dens.df.reg.2c <- ocean_region_lab(dens.df.reg.2c)
dens.df.reg.2c <- era_lab(dens.df.reg.2c, data_master=chum, breakpoint2=2011)
dens.df.reg.2c <- dplyr::filter(dens.df.reg.2c, !(Ocean.Region2 %in% c("BS", "GOA", "SEAK") & era=="Late"))



## Figures
ggplot(dens.df.st.2c) + 
  geom_vline(xintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_path(aes(x=x, y=dens, group=stock, col=ocean_region_lab), alpha=0.2) + 
  geom_path(data=dens.df.reg.2c, aes(x=x, y=dens, col=ocean_region_lab), alpha=0.85, linewidth=1) +
  scale_colour_manual(values=col.region) +
  facet_grid(rows=vars(era_lab), cols=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  coord_cartesian(xlim=c(-50,100)) +
  theme_sleek() + labs(x="Percent change in R/S", y="", col="Ocean Region") +
  theme(axis.text.y=element_blank(),
        legend.key.size = unit(10, "pt"),
        legend.position=c(0.9, 0.2),
        legend.background = element_blank(),
        legend.text = element_text(size = 8),
        legend.title = element_text(size=9),
        panel.spacing.y = unit(-0.5, "pt"),
        axis.ticks.y=element_blank()) 

# probably worth combining Chum-SEAK with another region in future analyses

```
**Figure 10**. Posterior probability distributions of the predicted effect of SST and competitors on Chum salmon productivity over three pre-defined time periods/eras (earliest in top panel). Regional mean effects are shown by bold lines and individual stocks’ distributions by light lines. Note that data from Alaska for the most recent time period are needed, and data in the Southeast Alaska region are sparse across all time periods (Figure 4).


## Random Walk models {.tabset}

### Sockeye
```{r sockeye random walk, warning=FALSE}
load("../output/models/dyn/sockeye/hbm_dyn_2c_sub.RData")


# Stock specific dataframe
probs <- c(0.025, 0.10, 0.50, 0.90, 0.975)
summ <- rstan::summary(dyn.2c.sub, pars = c("gamma", "kappa"), probs = probs)$summary
df.dyn.st.2c <- data.frame(Stock = sock$Stock,
                   Ocean.Region2 = sock$Ocean.Region2,
                   BY = sock$BY,
                   mu = summ[, "mean"],
                   se = summ[, "se_mean"],
                   lower_10 = summ[, "10%"],
                   upper_90 = summ[ , "90%"],
                   var = str_extract(rownames(summ), "[a-z]+"),
                   varnam = case_when(grepl("^gamma", rownames(summ)) ~ "SST",
                                      grepl("^kappa", rownames(summ)) ~ "Competitors")
                   )
df.dyn.st.2c <- ocean_region_lab(df.dyn.st.2c)

# Summarize at regional lvl
df.dyn.reg.2c <- dplyr::summarize(df.dyn.st.2c, 
                           reg_mean=mean(mu, na.rm=T), 
                           n_stk=n_distinct(Stock),
                           lower_10=quantile(mu, 0.1), 
                           upper_90=quantile(mu, 0.9), 
                           .by=c(Ocean.Region2, BY, varnam))
df.dyn.reg.2c <- ddply(df.dyn.reg.2c, .(Ocean.Region2), dplyr::filter, n_stk >= 4)
df.dyn.reg.2c <- ocean_region_lab(df.dyn.reg.2c)


# Figures
ggplot(df.dyn.reg.2c) +
  geom_hline(yintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +  
  geom_line(data= df.dyn.st.2c, aes(x=BY, y=mu, group=Stock, col=ocean_region_lab), alpha=0.2) +
  geom_line(aes(x=BY, y=reg_mean, col=ocean_region_lab), linewidth=1) +
  geom_ribbon(aes(x=BY, y=reg_mean, ymin=lower_10, ymax=upper_90, fill=ocean_region_lab), alpha=0.2) +
  facet_grid(cols=vars(factor(ocean_region_lab, levels=rev(levels(ocean_region_lab)))), rows=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  scale_y_continuous(limits=c(-1, 1), breaks=seq(-1,1,0.5)) + 
  scale_colour_manual(values=col.region, aesthetics=c("colour", "fill")) +
  theme_sleek() + theme(legend.position="none") +
   labs(x="Brood Year", y="Mean covariate effects")

# take out the scale squish, and revise 4-stock rule
```
**Figure 11**. Time-varying posterior mean estimates of SST and Competitor covariate effects on Sockeye productivity, modelled as a random walk. Individual stock estimates are in faint lines, while regional means and 80% CI are represented by bold lines and shaded areas. Region-wide means and CI are post-hoc calculations, rather than resulting from hierarchical model structures as in the stationary and era models.

### Pink
```{r pink random walk, warning=FALSE}
load("../output/models/dyn/pink/hbm_dyn_2c.RData")

# Stock specific dataframe
probs <- c(0.025, 0.10, 0.50, 0.90, 0.975)
summ <- rstan::summary(dyn.2c, pars = c("gamma", "kappa"), probs = probs)$summary
df.dyn.st.2c <- data.frame(Stock = pink$Stock,
                   Ocean.Region2 = pink$Ocean.Region2,
                   BY = pink$BY,
                   mu = summ[, "mean"],
                   se = summ[, "se_mean"],
                   lower_10 = summ[, "10%"],
                   upper_90 = summ[ , "90%"],
                   var = str_extract(rownames(summ), "[a-z]+"),
                   varnam = case_when(grepl("^gamma", rownames(summ)) ~ "SST",
                                      grepl("^kappa", rownames(summ)) ~ "Competitors"),
                   even_odd = ifelse(grepl("Even$", pink$Stock), "Even", "Odd")
                   )
df.dyn.st.2c <- ocean_region_lab(df.dyn.st.2c)

# Summarize at regional lvl
df.dyn.reg.2c <- dplyr::summarize(df.dyn.st.2c,
                                  reg_mean=mean(mu, na.rm=T), 
                                  n_stk=n_distinct(Stock),
                                  lower_10=quantile(mu, 0.1), 
                                  upper_90=quantile(mu, 0.9), 
                                  .by=c(Ocean.Region2, BY, varnam, even_odd))
df.dyn.reg.2c <- ddply(df.dyn.reg.2c, .(Ocean.Region2), dplyr::filter, n_stk >= 4)
df.dyn.reg.2c <- ocean_region_lab(df.dyn.reg.2c)


# Figures
ggplot(df.dyn.reg.2c) +
  geom_hline(yintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +  
  geom_line(data= df.dyn.st.2c, aes(x=BY, y=mu, group=Stock, col=ocean_region_lab), alpha=0.2) +
  geom_line(aes(x=BY, y=reg_mean, col=ocean_region_lab, linetype=even_odd), linewidth=1) +
  geom_ribbon(aes(x=BY, y=reg_mean, ymin=lower_10, ymax=upper_90, fill=ocean_region_lab, group=even_odd), alpha=0.2) +
  geom_line(data=dplyr::filter(df.dyn.st.2c, Ocean.Region2=="WC" & even_odd=="Even"), aes(x=BY, y=mu, group=Stock, col=ocean_region_lab), linetype="twodash", alpha=0.7) +
  facet_grid(cols=vars(factor(ocean_region_lab, levels=rev(levels(ocean_region_lab)))), rows=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  scale_linetype_manual(values=c("Odd" = "solid", "Even" = "twodash")) +
  scale_y_continuous(limits=c(-1,1.5), breaks=seq(-1,1.5,0.5)) + 
  scale_colour_manual(values=col.region, aesthetics=c("colour", "fill")) +
  theme_sleek() + theme(legend.position="none") +
   labs(x="Brood Year", y="Mean covariate effects")

```
**Figure 12**. Time-varying posterior mean estimates of SST and Competitor covariate effects on Pink salmon productivity, modelled as a random walk. Individual stock estimates are in faint lines, while regional means and 80% CI are represented by bold lines and shaded areas. Solid bold lines represent odd-year Pink stocks, while dashed lines are even-year stocks. Region-wide means and CI are post-hoc calculations, rather than resulting from hierarchical model structures as in the stationary and era models.

### Chum
```{r chum random walk, warning=FALSE}
load("../output/models/dyn/chum/hbm_dyn_2c.RData")


# Stock specific dataframe
probs <- c(0.025, 0.10, 0.50, 0.90, 0.975)
summ <- rstan::summary(dyn.2c, pars = c("gamma", "kappa"), probs = probs)$summary
df.dyn.st.2c <- data.frame(Stock = chum$Stock,
                   Ocean.Region2 = chum$Ocean.Region2,
                   BY = chum$BY,
                   mu = summ[, "mean"],
                   se = summ[, "se_mean"],
                   lower_10 = summ[, "10%"],
                   upper_90 = summ[ , "90%"],
                   var = str_extract(rownames(summ), "[a-z]+"),
                   varnam = case_when(grepl("^gamma", rownames(summ)) ~ "SST",
                                      grepl("^kappa", rownames(summ)) ~ "Competitors")
                   )
df.dyn.st.2c <- ocean_region_lab(df.dyn.st.2c)

# Summarize at regional lvl
df.dyn.reg.2c <- dplyr::summarize(df.dyn.st.2c, 
                           reg_mean=mean(mu, na.rm=T), 
                           n_stk=n_distinct(Stock),
                           lower_10=quantile(mu, 0.1), 
                           upper_90=quantile(mu, 0.9), 
                           .by=c(Ocean.Region2, BY, varnam))
df.dyn.reg.2c <- ddply(df.dyn.reg.2c, .(Ocean.Region2), dplyr::filter, n_stk >= 4)
df.dyn.reg.2c <- ocean_region_lab(df.dyn.reg.2c)


# Figures
ggplot(df.dyn.reg.2c[df.dyn.reg.2c$Ocean.Region2 != "SEAK",]) +
  geom_hline(yintercept = 0, color = "grey50", linetype = 2, linewidth = 0.25) +  
  geom_line(data= df.dyn.st.2c, aes(x=BY, y=mu, group=Stock, col=ocean_region_lab), alpha=0.2) +
  geom_line(aes(x=BY, y=reg_mean, col=ocean_region_lab), linewidth=1) +
  geom_ribbon(aes(x=BY, y=reg_mean, ymin=lower_10, ymax=upper_90, fill=ocean_region_lab), alpha=0.2) +
  facet_grid(cols=vars(factor(ocean_region_lab, levels=rev(levels(ocean_region_lab)))), rows=vars(factor(varnam, levels=c("SST", "Competitors")))) + 
  scale_y_continuous(limits=c(-1, 1), breaks=seq(-1,1,0.5)) + 
  scale_colour_manual(values=col.region, aesthetics=c("colour", "fill")) +
  theme_sleek() + theme(legend.position="none") +
   labs(x="Brood Year", y="Mean covariate effects")


```
**Figure 13**. Time-varying posterior mean estimates of SST and Competitor covariate effects on Chum salmon productivity, modelled as a random walk. Individual stock estimates are in faint lines, while regional means and 80% CI are represented by bold lines and shaded areas. Region-wide means and CI are post-hoc calculations, rather than resulting from hierarchical model structures as in the stationary and era models.


# Next steps

Priorities for future analysis are:

1. Update data to present (as recent as possible) for all species and regions, particularly Alaskan pink and chum.

2. Sensitivity analyses of SST and competitor index covariates

  _SST_
  
    The resolution of the SST data is currently 2 x 2 degrees ([ERSST](https://www.ncei.noaa.gov/products/extended-reconstructed-sst)). Cells within 400 km of ocean entry points are averaged over the three months after migration to sea (April-July, depending on location). Sensitivity to these assumptions will be tested with alternative spatial and temporal windows for averaging SST. However, these will remain consistent with the general hypothesis that juvenile salmon are impacted by temperature in the first year of marine life, and fine-scale analyses of local SST conditions are outside the scope of this project.
    There may be high-resolution SST data available in the near future to test.

 _Competitor Index_
 
    Our base cases analysis focuses on the abundance of pink salmon across the North Pacific in the second and/or third years of marine life as an index of potential direct or indirect competition for food. This approach is consistent with research that has suggested some salmon species primarily exhibit responses to competition with pink salmon during their second and/or third growing seasons at sea (Connors et al. 2020; Ruggerone et al.2023). However, we also plan to fit models to a small subset of alternative competitor indices as part of sensitivity analysis for this paper (e.g., total pink, chum and sockeye abundance; just North American abundances, biomass instead of abundance, etc.) and have begun sketching out a more comprehensive analytical plan for a separate paper that explores evidence for a wider range of hypotheses about intra- and inter-specific interactions at a variety of spatial and temporal scales. 


3. Random Walk model improvements
Work on versions of random walk models with (1) estimated region-wide mean effects, and (2) reduced sensitivity to time series of different lengths, is ongoing. 



```{r SST anomaly plot, include=FALSE}
## IGNORED for now

## Played with versions of SST plots to show marine heatwave period but all looked bad and left me with more questions than answers

sst.dat <- read.csv("../data/sst_yr_1_stock_anomalies.csv")
sst.raw <- read.csv("../data/sst_raw_anomalies.csv")
  
sst.dat.rm <- sst.dat[!duplicated(sst.dat[,c("Year", "sst_anomaly")]),] # removing duplicates in a hack way

# a better way:
all.info <- rbind(sock.info, pink.info, chum.info)
all.info <- all.info %>% dplyr::mutate(species = case_when(
      Stock.ID %in% seq(range(sock.info$Stock.ID)[1], range(sock.info$Stock.ID)[2]) ~ "sockeye",
      Stock.ID %in% seq(range(chum.info$Stock.ID)[1], range(chum.info$Stock.ID)[2]) ~ "chum",
      Stock.ID %in% seq(range(pink.info$Stock.ID)[1], range(pink.info$Stock.ID)[2]) ~ "pink")) %>%
  mutate(Stock = paste(Stock, species, sep="-"))
all.info.geo <- geographic.order(all.info)
all.info.geo$geo.id <- 1:nrow(all.info.geo)

sst.join <- left_join(sst.dat, all.info.geo[,c("Stock.ID", "lat", "lon", "geo.id")], by="Stock.ID")
sst.distinct <- sst.join[!duplicated(sst.join[,c("Year", "lat", "lon")]),]
sst.distinct <- left_join(sst.distinct, all.info[,c("ocean_region_lab", "Stock.ID")], by="Stock.ID")
# Plotting monthly data instead would be interesting

ggplot(sst.distinct) + geom_line(aes(x=Year, y=sst_anomaly, group=Stock.ID, col=geo.id), alpha=0.2) +
  geom_hline(yintercept=0, linetype="dashed") + 
  geom_rect(aes(xmin=2013, xmax=2015, ymin=-3, ymax=3), alpha=0.005, fill="orangered4") +
  coord_cartesian(xlim=c(2000, 2022), ylim=c(-3,3)) +
  viridis::scale_colour_viridis() +
  theme_sleek() 

# or use region colours but this just looks purple
ggplot(sst.distinct) + 
   geom_rect(aes(xmin=2012, xmax=2020, ymin=-3, ymax=3), fill="grey80") +
   geom_line(aes(x=Year, y=sst_anomaly, group=Stock.ID, col=ocean_region_lab), alpha=0.2) +
  geom_hline(yintercept=0, linetype="dashed") + 
  coord_cartesian(xlim=c(1960, 2022), ylim=c(-3,3)) +
  scale_colour_manual(values=col.region) +
  theme_sleek()  +
  facet_wrap(vars(ocean_region_lab))

# Or use the one we had before (our actual SST index)

# first get the data
all.brood <- bind_rows(fill.time.series(sock), fill.time.series(chum), fill.time.series(pink))
covar.dat.st <- all.brood %>% select(Stock, BY, Species, Ocean.Region2, early_sst, np_pinks_sec) %>% tidyr::pivot_longer(cols=c(early_sst), names_to = "covar") %>% mutate(covar_nam = "SST Index")
covar.dat.st <- ocean_region_lab(covar.dat.st)
covar.dat.reg <- dplyr::summarize(.data=covar.dat.st, mean_covar = mean(value), .by = c("BY", "ocean_region_lab", "covar_nam"))

ggplot(covar.dat.st) + 
  #geom_vline(xintercept=c(1976,1988), color = "grey50", linetype = 2, linewidth = 0.25) +
  geom_line(aes(x=BY, y=value, group=Stock, colour=ocean_region_lab), linewidth=0.5, alpha=0.1) +
  #geom_line(data=covar.dat.reg, aes(x=BY, y=mean_covar, col=ocean_region_lab), linewidth=0.5) +
  facet_grid(rows=NULL, cols=vars(ocean_region_lab), scales="free_y") +
  geom_rect(aes(xmin=2013, xmax=2015, ymin=-3, ymax=3), alpha=0.005, fill="orangered4") +
  scale_colour_manual(values=col.region) +
  coord_cartesian(xlim=c(2000, 2020)) +
  theme_sleek() + theme(legend.position="none") +
  labs(x= "Brood Year", y="")
# the above not very interesting at the moment but maybe to show marine heatwave windows at some point in future...

```
